---
title: "First Modelling of mClinica-Data"
output: html_document
---

```{r}

###### Aggregate Prices ######

aggregatePrices_SkuCode <- function(dataTrain) {
  
  price_names_prefix <- c("mean", "max", "min")
  price_names_skuCode <- paste0(price_names_prefix, "Price", "SkuCode", sep = "")
  price_names_skuPrefix <- paste0(price_names_prefix, "Price", "SkuPrefix", sep = "")
  
  lagDemand_names_skuCode <- paste0("demand_lag", 1:4, "_SkuCode", sep = "")
  lagDemand_names_skuPrefix <- paste0("demand_lag", 1:4, "_SkuPrefix", sep = "")
  
  lagCumDemand_names_skuCode <- paste0("cumDemand_lag", 1:4, "_SkuCode", sep = "")
  lagCumDemand_names_skuPrefix <- paste0("cumDemand_lag", 1:4, "_SkuPrefix", sep = "")
  
  staticCols <- c("sku_code_prefix", 
                  "quantity_perPackage",
                  price_names_skuCode,
                  price_names_skuPrefix,
                  lagDemand_names_skuCode,
                  lagDemand_names_skuPrefix,
                  lagCumDemand_names_skuCode,
                  lagCumDemand_names_skuPrefix,
                  "pharmaciesActive_month", 
                  "pharmaciesActive_sinceBeginning")
  
  dataTrain_agg <- dataTrain %>%
    group_by(sku_code) %>%
    group_by(Ordered_date_yearweek, add = TRUE) %>%
    group_by(.dots = staticCols, add = TRUE) %>%
    summarise(demand = sum(demand))
  
  return(as.data.frame(dataTrain_agg))
}

#---

####### Training Skus seperately #############

mClinicaML_split <- function(data = trainData,
                             model,
                             formula,
                             n.ahead = 4) {
  
  maxWeek <- max(data$Ordered_date_yearweek)
  
  dataGroups <- data %>%
    group_by(sku_code_prefix) %>%
    group_split()
  
  registerDoParallel()
  
  resultsPerSku <- foreach(dataGroup = dataGroups, .packages = c("dplyr", "caret")) %dopar% {
    
    if("molecule" %in% colnames(dataGroup)) {
      dataGroup <- dataGroup %>%
        select(-molecule)
    }
    
    dataTrain <- dataGroup %>%
      filter(Ordered_date_yearweek <= maxWeek - n.ahead)
    
    dataTest <- setdiff(dataGroup, dataTrain)
    
    dataTrain_mod <- dataTrain %>%
      select(-c(sku_code, sku_code_prefix))
    
    dataTest_mod <- dataTest %>%
      select(-c(sku_code, sku_code_prefix))
    
    # 'Caret' bietet die Möglichkeit automatisiertes Parametertuning durchführen zu
    # können (sofern das Modell tune-bare Parameter besitzt).
    fitControl <- trainControl(method = "cv",
                               number = 5)
    
    fit <- train(formula, 
                 data = dataTrain_mod, 
                 method = model, 
                 trControl = fitControl)
    
    preds <- predict(fit, newdata = dataTest_mod)
    
    dataPred <- cbind(dataTest, preds)
    
    dataPred_agg <- dataPred %>%
      group_by(sku_code, Ordered_date_yearweek) %>%
      summarise("demand" = sum(demand), "preds" = sum(preds))
    
    return(list(fit = fit, 
                preds = preds, 
                actuals = dataTest$demand,
                preds_agg = dataPred_agg$preds,
                actuals_agg = dataPred_agg$demand,
                "sku_code_prefix" = unique(dataGroup$sku_code_prefix)))
  }
  
  sku_codes_prefix <- sapply(resultsPerSku, function(result) {
    result$sku_code_prefix
  })
  names(resultsPerSku) <- sku_codes_prefix
  
  fits <- lapply(resultsPerSku, function(result) {
    result$fit
  })
  names(fits) <- sku_codes_prefix
  
  lengths <- sapply(resultsPerSku, function(result) {
    length(result$preds)
  })
  maxLength <- max(lengths)
  
  # Evaluating Results
  predsMatrix <- sapply(resultsPerSku, function(result) {
    preds <- result$preds
    if(length(preds) < maxLength) {
      preds <- c(preds, rep.int(NA, times = maxLength - length(preds)))
      names(preds) <- 1:length(preds) 
    }
    
    return(preds)
  })
  
  actualsMatrix <- sapply(resultsPerSku, function(result) {
    actuals <- result$actuals
    
    if(length(actuals) < maxLength) {
      actuals <- c(actuals, rep.int(NA, times = maxLength - length(actuals)))
      names(actuals) <- 1:length(actuals)
    }
    
    return(actuals)
  })
  
  errorsMatrix <- actualsMatrix - predsMatrix
  
  errorMeasures_calc <- sapply(1:ncol(errorsMatrix), function(j) {
    preds <- predsMatrix[, j]
    actuals <- actualsMatrix[, j]
    errors <- errorsMatrix[, j]
    
    rmse <- rmse(errors)
    mae <- mae(errors)
    smape <- smape(actuals, 
                   preds)
    
    return(c(rmse, mae, smape))
  })
  errorMeasures_calc <- t(errorMeasures_calc)
  colnames(errorMeasures_calc) <- c("rmse", "mae", "smape")
  
  return(list("errorMeasurement" = errorMeasures_calc,
              "preds" = predsMatrix,
              "actuals" = actualsMatrix,
              "errors" = errorsMatrix,
              "fits" = fits))
}

############# Training Skus simultanously ##############

mClinicaML_noSplit <- function(data = trainData,
                               model,
                               formula,
                               n.ahead = 4) {
  
  if("molecule" %in% colnames(data)) {
    data <- data %>%
      select(-molecule)
  }
  
  maxWeek <- max(data$Ordered_date_yearweek)
  
  dataTrain <- data %>%
    filter(Ordered_date_yearweek <= maxWeek - n.ahead)
  
  dataTest <- setdiff(data, dataTrain)
  
  dataTrain_mod <- dataTrain %>%
    select(-c(sku_code, sku_code_prefix))
  
  dataTest_mod <- dataTest %>%
    select(-c(sku_code, sku_code_prefix))
  
  # 'Caret' bietet die Möglichkeit automatisiertes Parametertuning durchführen zu
  # können (sofern das Modell tune-bare Parameter besitzt).
  fitControl <- trainControl(method = "cv",
                             number = 5)
  
  fit <- train(formula, 
               data = dataTrain_mod, 
               method = model, 
               trControl = fitControl)
  
  preds <- predict(fit, newdata = dataTest_mod)
  actuals <- dataTest$demand
  errors <- actuals - preds
  
  rmse <- rmse(errors)
  mae <- mae(errors)
  smape <- smape(actuals, 
                 preds)
  
  errorMeasurement <- c(rmse, mae, smape)
  names(errorMeasurement) <- c("rmse", "mae", "smape") 
  
  dataPred <- cbind(dataTest, preds)
  
  dataPred_agg <- dataPred %>%
    group_by(sku_code, Ordered_date_yearweek) %>%
    summarise("demand" = sum(demand), "preds" = sum(preds))
  
  preds_agg <- dataPred_agg$preds
  actuals_agg <- dataPred_agg$demand
  errors_agg <- actuals_agg - preds_agg
  
  rmse_agg <- rmse(errors_agg)
  mae_agg <- mae(errors_agg)
  smape_agg <- smape(actuals_agg, 
                     preds_agg)
  
  errorMeasurement_agg <- c(rmse_agg, mae_agg, smape_agg)
  names(errorMeasurement_agg) <- c("rmse", "mae", "smape") 
  
  return(list("errorMeasurement" = errorMeasurement,
              "preds" = preds,
              "actuals" = actuals,
              "preds_agg" = dataPred_agg$preds,
              "actuals_agg" = dataPred_agg$demand,
              "errorMeasurement_agg" = errorMeasurement_agg,
              "errors" = errors,
              "fit" = fit))
}

#---

######## Over/Under Counts ########

# The argument 'errors' can either be a vector or a matrix. If a matrix is provided the errors of each model
# have to be 
getOverUnderPercent <- function(errors) {
  
  if(is.null(nrow(errors))) {
    errors <- errors[!is.na(errors)]
    overPercent <- sum(errors < 0) / length(errors)
    underPercent <- sum(errors > 0) / length(errors)
    
    return(data.frame(overPercent, underPercent))
    
  } else {
    
    overUnderPercent <- apply(errors, MARGIN = 2, function(errors_col) {
      errors_col <- errors_col[!is.na(errors_col)]
      overPercent <- sum(errors_col < 0) / length(errors_col)
      underPercent <- sum(errors_col > 0) / length(errors_col)
      return(c(overPercent, underPercent))
    })
    rownames(overUnderPercent) <- c("overPercent", "underPercent")
    return(overUnderPercent)
  }
}

#---

######## Plot Preds ########

plotPreds <- function(preds,
                      actuals) {
  
  preds <- preds[!is.na(preds)]
  actuals <- actuals[!is.na(actuals)]
  
  range_model <- round(max(preds) - min(preds), 2)
  range_real  <- round(max(actuals) - min(actuals), 2)
  
  plotDataActuals <- data.frame("number" = seq_along(preds), 
                                "demand" = actuals, 
                                label = "actuals")
  
  plotDataPreds <- data.frame("number" = seq_along(actuals), 
                              "demand" = preds, 
                              label = "preds")
  
  plotData <- rbind(plotDataActuals, plotDataPreds)
  
  g <- ggplot(data = plotData, aes(x = number, y = demand, col = label, group = number)) +
    geom_point() + 
    geom_line() + 
    labs(x = "Combination of week, sku_code and price", 
         y = "demand", 
         title = paste("Predicted vs. Actual Demand"),
         subtitle = paste("range actuals vs. range predicted Values:", range_real, "vs.", range_model)) + 
    theme_bw()
  
  print(g)
}

#---

######## Boxplot Errors #########

# errors can either be a matrix of errors, whose columns belong to one model or can be a vector of errors all
# belonging to the same model.

boxplotErrors <- function(errors,
                          abs = FALSE) {
  
  if(is.null(nrow(errors))) {
    errors <- as.data.frame(errors)
  }

  if(abs) {
    errors <- abs(errors)
  }
  
  ranges <- sapply(1:ncol(errors), function(j) {
    errors_col <- errors[, j]
    errors_col <- errors_col[!is.na(errors_col)]
    min <- min(errors_col)
    max <- max(errors_col)
    return(c(min, max))
  })
  
  minRange <- min(ranges[1, ])
  maxRange <- max(ranges[2, ])
  
  errors_list <- lapply(1:ncol(errors), function(j) {
    errors_col <- errors[, j]
    return(errors_col[!is.na(errors_col)])
  })
  names(errors_list) <- colnames(errors)
  
  ggplot(stack(errors_list), aes(x = ind, y = values)) +
    geom_boxplot(fill = "steelblue") + 
    coord_cartesian(ylim=c(minRange, maxRange)) +
    labs(x = "Model", 
         y = "Error", 
         title = "Distribution of errors among all fitted models") + 
    theme_bw()
}

######## Error Measures #########

mse <- function(errors) {
  mse <- mean((errors)^2, na.rm = TRUE)
  
  mse <- round(mse, digits = 3)
  return(mse)
}

rmse <- function(errors) {
  rmse <- sqrt(mean((errors)^2, na.rm = TRUE))
  
  rmse <- round(rmse, digits = 3)
  return(rmse)
}

mae <- function(errors) {
  mae <- mean(abs(errors), na.rm = TRUE)
  
  mae <- round(mae, digits = 3)
  return(mae)
}

smape <- function(actuals,
                  preds) {
  
  if(any(abs(actuals[!is.na(actuals)]) + abs(preds[!is.na(preds)]) == 0)) {
    index_zero <- which(abs(actuals) + abs(preds) == 0)
    actuals[index_zero] <- 1
  }
  
  errors <- actuals - preds
  smape <- mean(abs(errors) / (abs(actuals) + abs(preds) / 2), na.rm = TRUE)
  
  smape <- round(smape, digits = 3)
  return(smape)
}

getTopSkus <- function(transactions,
                       rankingByPrefix = TRUE,
                       topPercentSkus = 0.0005) {
  
  if(rankingByPrefix) sku_feature <- "sku_code_prefix" else sku_feature <- "sku_code"
  
  countPerSku <- transactions %>%
    group_by(.dots = sku_feature) %>%
    summarise(count = n())
  
  sku_codes_top <- countPerSku %>%
    filter(count >= quantile(countPerSku$count, 1 - topPercentSkus))
  
  rankedSkus <- sku_codes_top[order(sku_codes_top$count, decreasing = TRUE), ]
  rankedSkus <- c(rankedSkus[[sku_feature]])
  
  # If we have been working with sku_codes_prefixes we will grab all sku_codes that belong to the
  # determined top sku_code_prefixes. 
  if(!rankingByPrefix) {
    return(rankedSkus) 
    
  } else {
    sku_codes_available <- unique(transactions$sku_code)
    sku_codes_prefix_available <- str_split(sku_codes_available, pattern = "ID|-", simplify = TRUE)[, 2]
    topSkuCodes <- sku_codes_available[sku_codes_prefix_available %in% rankedSkus]
    
    return(topSkuCodes)
  }
} 


```

## Laden der benötigten Packages
```{r, echo = TRUE, message = FALSE}
library(tidyverse)
library(caret)
library(mlbench)
library(modelr)
library(gtools)
library(Gmisc)
library(smooth)
library(doParallel)
library(foreach)
```

# DATA 1 - Only skuCodes, one Price, no extrapolation
1) We start of with the simplest kind of training data that is avaiable for us. We work on a weekly skuCode-level basis with only one price observation per week. 
2) For now we only work with the top 1 percent of Skus.
3) No extrapolation of prices took place here.

## Splitting between sku_code_prefixes
For the seperated fitting of every sku_code_prefix we work only with the top 0,5% of sku_codes to save computation time.
```{r}
transactions <- readRDS("transactions_clean.rds")
dataTrain <- readRDS("training_withoutDistr_noExtrapolation.rds")
sku_codes_top <- getTopSkus(transactions, rankingByPrefix = TRUE, topPercentSkus = 0.005)

dataTrain_top <- dataTrain %>%
  filter(sku_code %in% sku_codes_top)

dataTrain_noDistr_noExtra_onePrice <- aggregatePrices_SkuCode(dataTrain_top)

xgbTree_split_onePrice <- mClinicaML_split(data = dataTrain_noDistr_noExtra_onePrice, 
                                           model = "xgbTree", 
                                           formula = demand~., 
                                           n.ahead = 4)
```

```{r}
errorMeasurement <- xgbTree_split_onePrice$errorMeasurement
summary(errorMeasurement)
```

```{r}
overUnderPercent <- getOverUnderPercent(xgbTree_split_onePrice$errors)
summary(overUnderPercent[1, ])
```

```{r}
boxplotErrors(xgbTree_split_onePrice$errors)
```

```{r}
sku_code_prefix <- "102068"
preds_sku <- xgbTree_split_onePrice$preds[, sku_code_prefix]
actuals_sku <- xgbTree_split_onePrice$actuals[, sku_code_prefix]

plotPreds(preds = preds_sku,
         actuals = actuals_sku)

```

## No splitting between sku_code_prefixes
For the none-splitted fitting we work with the top 1% of sku_codes_prefixes.
```{r}
sku_codes_top <- getTopSkus(transactions, rankingByPrefix = TRUE, topPercentSkus = 0.01)

dataTrain_top <- dataTrain %>%
  filter(sku_code %in% sku_codes_top)

dataTrain_noDistr_noExtra_onePrice <- aggregatePrices_SkuCode(dataTrain_top)

xbgTree_noSplit_onePrice <- mClinicaML_noSplit(data = dataTrain_noDistr_noExtra_onePrice, 
                                             model = "xgbTree", 
                                             formula = demand~., 
                                             n.ahead = 4)
xgbTree_noSplit_onePrice <- xbgTree_noSplit_onePrice
```

```{r}
xgbTree_noSplit_onePrice$errorMeasurement
```

```{r}
overUnderPercent <- getOverUnderPercent(xgbTree_noSplit_onePrice$errors)
overUnderPercent
```

```{r}
boxplotErrors(xgbTree_noSplit_onePrice$errors)
```

```{r}
preds_sku <- xgbTree_noSplit_onePrice$preds_agg
actuals_sku <- xgbTree_noSplit_onePrice$actuals_agg

plotPreds(preds = preds_sku,
         actuals = actuals_sku)
```

# Data 2 - Only SkuCodes, no extrapolation

1) We try to work with extrapolated price data now.
2) For now we work on sku_code-level and disregard distributors
3) Yet again we only work with the top 1% of sku_code_prefixes

```{r}
dataTrain <- readRDS("training_withoutDistr_noExtrapolation.rds")
sku_codes_top <- getTopSkus(transactions, rankingByPrefix = TRUE, topPercentSkus = 0.01)

dataTrain_noDistr_noExtra <- dataTrain %>%
  filter(sku_code %in% sku_codes_top)

xgbTree_noSplit_noDistr_noExtra <- mClinicaML_noSplit(data = dataTrain_noDistr_noExtra, 
                                                    model = "xgbTree", 
                                                    formula = demand~., 
                                                    n.ahead = 4)
```

```{r}
xgbTree_noSplit_noDistr_noExtra$errorMeasurement
```

```{r}
overUnderPercent <- getOverUnderPercent(xgbTree_noSplit_noDistr_noExtra$errors)
overUnderPercent
```

```{r}
boxplotErrors(xgbTree_noSplit_noDistr_noExtra$errors)
```

```{r}
preds_sku <- xgbTree_noSplit_noDistr_noExtra$preds_agg
actuals_sku <- xgbTree_noSplit_noDistr_noExtra$actuals_agg

plotPreds(preds = preds_sku,
         actuals = actuals_sku)
```

# Data 3 - Only SkuCodes, extrapolation

```{r}
dataTrain <- readRDS("training_withoutDistr_extrapolated.rds")
sku_codes_top <- getTopSkus(transactions, rankingByPrefix = TRUE, topPercentSkus = 0.01)

dataTrain_noDistr_extra <- dataTrain %>%
  filter(sku_code %in% sku_codes_top)

xgbTree_noDistr_extra <- mClinicaML_noSplit(data = dataTrain_noDistr_extra, 
                                            model = "xgbTree", 
                                            formula = demand~.,
                                            demandOfUnit = TRUE,
                                            n.ahead = 4)
```

```{r}
overUnderPercent <- getOverUnderPercent(xgbTree_noDistr_extra$res_noAgg$error)
overUnderPercent
```

```{r}
xgbTree_noDistr_extra$accuracy
```

```{r}
boxplotErrors(xgbTree_noDistr_extra$res_noAgg$error)
```

```{r}
res_aggSkuPrefix <- xgbTree_noDistr_extra$res_aggSkuPrefix
res_hasProm <- res_aggSkuPrefix[res_aggSkuPrefix$hasPromotion, ]

mae(res_hasProm$actual, res_hasProm$pred) / mean(res_hasProm$actual)

summary(xgbTree_noDistr_extra$res_noAgg$actual)
summary(xgbTree_noDistr_extra$res_noAgg$pred)
```

```{r}
res_filtered <- res_aggSkuPrefix %>%
  filter(sku_code_prefix == "102065")

preds <- res_filtered$pred
actuals <- res_filtered$actual

plotPreds(preds = preds,
         actuals = actuals)

res_filtered <- res_aggSkuPrefix %>%
  filter(sku_code_prefix == "102068")

preds <- res_filtered$pred
actuals <- res_filtered$actual

plotPreds(preds = preds,
         actuals = actuals)
```




# Data 4 - Only SkuCodes, no extrapolation

```{r}
dataTrain <- readRDS("training_withoutDistr_noExtrapolation.rds")
sku_codes_top <- getTopSkus(transactions, rankingByPrefix = TRUE, topPercentSkus = 0.01)

dataTrain_noDistr_noExtra <- dataTrain %>%
  filter(sku_code %in% sku_codes_top)

xgbTree_noDistr_noExtra <- mClinicaML_noSplit(data = dataTrain_noDistr_noExtra, 
                                            model = "xgbTree", 
                                            formula = demand~.,
                                            demandOfUnit = TRUE,
                                            n.ahead = 4)
```

```{r}
overUnderPercent <- getOverUnderPercent(xgbTree_noDistr_noExtra$res_noAgg$error)
overUnderPercent
```

```{r}
xgbTree_noDistr_noExtra$accuracy
```

```{r}
boxplotErrors(xgbTree_noDistr_noExtra$res_noAgg$error)
```

```{r}
res_aggSkuPrefix <- xgbTree_noDistr_noExtra$res_aggSkuPrefix
noZero_index <- res_aggSkuPrefix$actual != 0

res_aggSkuPrefix

mape(res_aggSkuPrefix$actual[noZero_index], res_aggSkuPrefix$pred[noZero_index])

res_hasProm <- res_aggSkuPrefix[res_aggSkuPrefix$hasPromotion, ]

mae(res_hasProm$actual, res_hasProm$pred) / mean(res_hasProm$actual)

summary(xgbTree_noDistr_noExtra$res_noAgg$actual)
summary(xgbTree_noDistr_noExtra$res_noAgg$pred)
```

```{r}
res_filtered <- res_aggSkuPrefix %>%
  filter(sku_code_prefix == "102065")

preds <- res_filtered$pred
actuals <- res_filtered$actual
hasPromotion <- res_filtered$hasPromotion

plotPreds(preds = preds,
         actuals = actuals,
         hasPromotion = hasPromotion)

res_filtered <- res_aggSkuPrefix %>%
  filter(sku_code_prefix == "102068")

preds <- res_filtered$pred
actuals <- res_filtered$actual
hasPromotion <- res_filtered$hasPromotion

plotPreds(preds = preds,
         actuals = actuals,
         hasPromotion = hasPromotion)


res_filtered <- res_aggSkuPrefix %>%
  filter(sku_code_prefix == "101942")

preds <- res_filtered$pred
actuals <- res_filtered$actual
hasPromotion <- res_filtered$hasPromotion

plotPreds(predData = res_aggSkuPrefix,
          sku_code_prefix = "101942")



```




